---
title: "Categories of CCC"
author: 
- Peter DeWitt
- James Feinstein
- Tell Bennett
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{categories-of-ccc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE)
```

# Introduction
This vignette describes how the `pccc` package generates the CCC categories from ICD-9 and
ICD-10 codes. The categorization is based on the work of Feudtner et al.(2014). A reference document
showing the lists of codes for each category was published as a supplement to Feudtner et al.(2014) and we
have made it available as part of the `pccc` package. After installing the package, you can find the file on your system
with the below `system.file` call. Open the file with your preferred/available program
for `.docx` files (Word, etc.).

```{r, eval = FALSE}
system.file("pccc_references/Categories_of_CCCv2_and_Corresponding_ICD.docx", package = "pccc")
```

To evaluate the code chunks in this example you will need to load the
following R packages.
```{r, message = FALSE}
library(pccc)
library(dplyr)
```

# Generating CCC categories from ICD codes
The `ccc` function is the workhorse here. Simply put, a user will provide ICD codes as strings
and `ccc` will return CCC categories. The ICD codes are matched exactly and should include leading zeroes,
but should not include periods. Users of the `pccc` package will need to pre-process the ICD-9
and ICD-10 codes in their data so that the strings are formatted in the way that
the `pccc` package will recognize them. For example, ICD-9-CM 04.92 will only be matched by the string "0492". 
"492" _will not_ match.  

The `data.frame` named `dat` below has information about three patients (A-C), each of whom has a different 
string for the ICD-9 code 04.92. The `ccc` function will only identify subject `A`
as having a CCC.

```{r}
dat <- data.frame(ids = c("A", "B", "C"), code = c("0492", "492", "04.92"))
dat
ccc(dat, ids, code, code, icdv = 9)
```

## Viewing ICD Codes
The ICD codes used to define the categories of CCC have been hard-coded into the
C++ backend of the `pccc` package.  The `get_codes` function provides the end user
a easy way to view all of the ICD codes as strings.

```{r}
get_codes(9)   # ICD-9
get_codes(10)  # ICD-10
```

`get_codes` returns a matrix of character vectors.  The
categories of CCC are the rows and the diagnostic (`dx`) or
procedure (`pc`) codes are the columns.  If you need to view the ICD-9 codes for `cvd`, for
example, subset as usual for matrices.
```{r}
get_codes(9)["cvd", ]
```

If a user prefers to explore the ICD codes in a `data.frame`, then use the base R
`as.data.frame` method to convert the output from `get_codes` to a
`data.frame`.

```{r}
as.data.frame(get_codes(9))
```

# Neuromuscular
## ICD-9
```{r}
get_codes(9)["neuromusc", ]
```

## ICD-10
```{r}
get_codes(10)["neuromusc", ]
```

# CVD
## ICD-9
```{r}
get_codes(9)["cvd", ]
```

## ICD-10
```{r}
get_codes(10)["cvd", ]
```

# Respiratory
## ICD-9
```{r}
get_codes(9)["respiratory", ]
```

## ICD-10
```{r}
get_codes(10)["respiratory", ]
```

# Renal
## ICD-9
```{r}
get_codes(9)["renal", ]
```

## ICD-10
```{r}
get_codes(10)["renal", ]
```

# GI
## ICD-9
```{r}
get_codes(9)["gi", ]
```

## ICD-10
```{r}
get_codes(10)["gi", ]
```

# Hemato Immu
## ICD-9
```{r}
get_codes(9)["hemato_immu", ]
```

## ICD-10
```{r}
get_codes(10)["hemato_immu", ]
```

# Metabolic
## ICD-9
```{r}
get_codes(9)["metabolic", ]
```

## ICD-10
```{r}
get_codes(10)["metabolic", ]
```

# Congeni Genetic
## ICD-9
```{r}
get_codes(9)["congeni_genetic", ]
```

## ICD-10
```{r}
get_codes(10)["congeni_genetic", ]
```

# Malignancy
## ICD-9
```{r}
get_codes(9)["malignancy", ]
```

## ICD-10
```{r}
get_codes(10)["malignancy", ]
```

# Neonatal
## ICD-9
```{r}
get_codes(9)["neonatal", ]
```

## ICD-10
```{r}
get_codes(10)["neonatal", ]
```

# Tech Dep
## ICD-9
```{r}
get_codes(9)["tech_dep", ]
```

## ICD-10
```{r}
get_codes(10)["tech_dep", ]
```

# Transplant
## ICD-9
```{r}
get_codes(9)["transplant", ]
```

## ICD-10
```{r}
get_codes(10)["transplant", ]
```

# References

* Feudtner C, et al. Pediatric complex chronic conditions classification system version 2: updated for ICD-10 and complex medical technology dependence and transplantation, BMC Pediatrics, 2014, 14:199, DOI: 10.1186/1471-2431-14-199
