---
title: "Categories of CCC"
author: "Peter DeWitt"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{categories-of-ccc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE)
```

This vignette reports the ICD codes, as formated for string matching, by
Categories of CCC.  A references document, provided as a supplement to Feudtner
2014, is provided with the pccc package.  You can find the file on your system
with a `system.file` call.  Open the file with your preferred/available program
for `.docx` files.

```{r, eval = FALSE}
system.file("pccc_references/Categories_of_CCCv2_and_Corresponding_ICD.docx", package = "pccc")
```

To evaluate the code chunks in this example you will need to attach the
following namespaces.
```{r, message = FALSE}
library(pccc)
library(dplyr)
```

# ICD Codes
The ICD codes used in this work are matched exactly.  For example, ICD-9 Code
04.92 for subcategory Devices under Category Neurologic or Neuromuscular will be
matched by the string 0492.  492 _will not_ match.  For example, the
`data.frame` `dat` below has three subjects, each will a different, and
reasonable, string for the ICD-9 code 04.92.  When sent to the `ccc` function we
only get a positive return for subject `A`.

```{r}
dat <- data.frame(ids = c("A", "B", "C"), code = c("0492", "492", "04.92"))
dat
ccc(dat, ids, vars(code), vars(code), icdv = 9)
```

It is critical that the end users of the pccc package understand how the ICD-9
and ICD-10 codes they have are coded and that the strings are as expected for
the pccc package.

## Viewing ICD Codes
The ICD codes used to define the categories of CCC have been hard coded into the
C++ backend of the pccc package.  The `get_codes` function provides the end user
a easy way to view all the strings.

```{r}
get_codes(9)   # ICD-9
get_codes(10)  # ICD-10
```

The codes are returned to the end user as a matrix of character vectors.  The
categories of CCC are the rows and the columns for the diagnostic (`dx`) or
procedure (`pc`) codes.  If you need to view the ICD-9 codes for `cvd`, for an
example, subset as usual for matrices.
```{r}
get_codes(9)["cvd", ]
```

If a `data.frame` would be more helpful for you to expore the ICD codes with
then use the `as.data.frame` method to convert the return from `get_codes` to a
`data.frame`.

```{r}
as.data.frame(get_codes(9))
```

# Neuromuscular
## ICD-9
```{r}
get_codes(9)["neuromusc", ]
```

## ICD-10
```{r}
get_codes(10)["neuromusc", ]
```

# CVD
## ICD-9
```{r}
get_codes(9)["cvd", ]
```

## ICD-10
```{r}
get_codes(10)["cvd", ]
```

# Respiratory
## ICD-9
```{r}
get_codes(9)["respiratory", ]
```

## ICD-10
```{r}
get_codes(10)["respiratory", ]
```

# Renal
## ICD-9
```{r}
get_codes(9)["renal", ]
```

## ICD-10
```{r}
get_codes(10)["renal", ]
```

# GI
## ICD-9
```{r}
get_codes(9)["gi", ]
```

## ICD-10
```{r}
get_codes(10)["gi", ]
```

# Hemato Immu
## ICD-9
```{r}
get_codes(9)["hemato_immu", ]
```

## ICD-10
```{r}
get_codes(10)["hemato_immu", ]
```

# Metabolic
## ICD-9
```{r}
get_codes(9)["metabolic", ]
```

## ICD-10
```{r}
get_codes(10)["metabolic", ]
```

# Congeni Genetic
## ICD-9
```{r}
get_codes(9)["congeni_genetic", ]
```

## ICD-10
```{r}
get_codes(10)["congeni_genetic", ]
```

# Malignancy
## ICD-9
```{r}
get_codes(9)["malignancy", ]
```

## ICD-10
```{r}
get_codes(10)["malignancy", ]
```

# Neonatal
## ICD-9
```{r}
get_codes(9)["neonatal", ]
```

## ICD-10
```{r}
get_codes(10)["neonatal", ]
```

# Tech Dep
## ICD-9
```{r}
get_codes(9)["tech_dep", ]
```

## ICD-10
```{r}
get_codes(10)["tech_dep", ]
```

# Transplant
## ICD-9
```{r}
get_codes(9)["transplant", ]
```

## ICD-10
```{r}
get_codes(10)["transplant", ]
```


# References

* Feudtner C, et al. Pediatric complex chronic conditions classification system version 2: updated for ICD-10 and complex medical technology dependence and transplantation, BMC Pediatrics, 2014, 14:199, DOI: 10.1186/1471-2431-14-199
