---
title: "Pediatric Complex Chronic Conditions"
author: 
- Peter DeWitt
- James Feinstein
- Tell Bennett
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pccc-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Introduction
This vignette describes how the `pccc` package generates the Complex Chronic Condition Categories (CCC) from ICD-9 and
ICD-10 codes. The categorization is based on the work of Feudtner et al.(2014). A reference document
showing the lists of codes for each category was published as a supplement to Feudtner et al.(2014) and we
have made it available as part of the `pccc` package. After installing the package, you can find the file on your system
with the below `system.file` call. Open the file with your preferred/available program
for `.docx` files (Word, etc.).

```{r, eval = FALSE}
system.file("pccc_references/Categories_of_CCCv2_and_Corresponding_ICD.docx", package = "pccc")
```

To evaluate the code chunks in this example you will need to load the
following R packages.

```{r, message = FALSE}
library(pccc)
library(dplyr)
```

# Generating CCC categories from ICD codes
The `ccc` function is the workhorse here. Simply put, a user will provide ICD codes as strings
and `ccc` will return CCC categories. The ICD codes are matched exactly and should include leading zeroes,
but should not include periods. Users of the `pccc` package will need to pre-process the ICD-9
and ICD-10 codes in their data so that the strings are formatted in the way that
the `pccc` package will recognize them. For example, ICD-9-CM 04.92 will only be matched by the string "0492". 
"492" _will not_ match.

Specific rules for ICD Codes are:

* Codes should be alphanumeric only - no periods, spaces or other separator characters (e.g. _Diabetes with renal manifestations, type II or unspecified type, uncontrolled_ should be sent as 25042)
* ICD 9 codes should be at a minimum 3 digits long:
    * Codes less than 10 should be left padded with 2 zeros. E.g. _Cholera due to vibrio cholerae el tor_ should be sent as 0011) 
    * Codes less than 100 should be left padded with 1 zero. E.g. _Whooping cough_ should be sent as 033)
    
Users of PCCC may find the R Package [ICD](https://jackwasey.github.io/icd/) useful.

# PCCC Example

The `data.frame` named `dat` below has information about three patients (A-C), each of whom has a different 
string for the ICD-9 code 04.92. The `ccc` function will only identify subject `A`
as having a CCC.

## Basic Example

```{r}
dat <- data.frame(ids = c("A", "B", "C"), code = c("0492", "492", "04.92"))
dat
ccc(dat, ids, code, code, icdv = 9)
```

## Full Featured Example

```{r}

# Seth to provide code here

```


# Logic employed

There are 12 total categories of CCCs used in this package. The first group of 10 are mutually exclusive - only of them can be derived from a ICD code:

* Neurologic and Neuromuscular
* Cardiovascular
* Respiratory
* Renal and Urologic
* Gastrointestinal
* Hematologic or immunologic
* Metabolic
* Other Congenital or Genetic Defect
* Malignancy
* Premature and Neonatal

The last 2 can be be selected in addition to the above codes - so for example, one ICD code could be both a Gastrointestinal and Technical Dependency:

* Technical Dependency
* Transplant

Most of the ICD diagnosis codes are matched by substring, so as new versions of ICD-10-CM are released, some refinement of the CCC code lists may be required.

To see actual specific ICD codes by category, see [pccc-icd-codes](pccc-icd-codes.html)

# References

* Feudtner C, et al. Pediatric complex chronic conditions classification system version 2: updated for ICD-10 and complex medical technology dependence and transplantation, BMC Pediatrics, 2014, 14:199, DOI: 10.1186/1471-2431-14-199
