---
title: "pccc-icd-codes"
author: 
- Peter DeWitt
- James Feinstein
- Tell Bennett
- Seth Russell
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pccc-icd-codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

```

## Viewing ICD Codes
The ICD codes used to define the categories of CCC have been hard-coded into the
C++ backend of the `pccc` package.  The `get_codes` function provides the end user
a easy way to view all of the ICD codes as strings.

Make sure the library is loaded first

```{r}
library(pccc)
```

Call get_codes function:

```{r}
get_codes(9)   # ICD-9
get_codes(10)  # ICD-10
```

`get_codes` returns a matrix of character vectors.  The
categories of CCC are the rows and the diagnostic (`dx`) or
procedure (`pc`) codes are the columns.  If you need to view the ICD-9 codes for `cvd`, for
example, subset as usual for matrices.
```{r}
get_codes(9)["cvd", ]
```

If a user prefers to explore the ICD codes in a `data.frame`, then use the base R
`as.data.frame` method to convert the output from `get_codes` to a
`data.frame`.

```{r}
as.data.frame(get_codes(9))
```

# Neuromuscular
## ICD-9
```{r}
get_codes(9)["neuromusc", ]
```

## ICD-10
```{r}
get_codes(10)["neuromusc", ]
```

# CVD
## ICD-9
```{r}
get_codes(9)["cvd", ]
```

## ICD-10
```{r}
get_codes(10)["cvd", ]
```

# Respiratory
## ICD-9
```{r}
get_codes(9)["respiratory", ]
```

## ICD-10
```{r}
get_codes(10)["respiratory", ]
```

# Renal
## ICD-9
```{r}
get_codes(9)["renal", ]
```

## ICD-10
```{r}
get_codes(10)["renal", ]
```

# GI
## ICD-9
```{r}
get_codes(9)["gi", ]
```

## ICD-10
```{r}
get_codes(10)["gi", ]
```

# Hemato Immu
## ICD-9
```{r}
get_codes(9)["hemato_immu", ]
```

## ICD-10
```{r}
get_codes(10)["hemato_immu", ]
```

# Metabolic
## ICD-9
```{r}
get_codes(9)["metabolic", ]
```

## ICD-10
```{r}
get_codes(10)["metabolic", ]
```

# Congeni Genetic
## ICD-9
```{r}
get_codes(9)["congeni_genetic", ]
```

## ICD-10
```{r}
get_codes(10)["congeni_genetic", ]
```

# Malignancy
## ICD-9
```{r}
get_codes(9)["malignancy", ]
```

## ICD-10
```{r}
get_codes(10)["malignancy", ]
```

# Neonatal
## ICD-9
```{r}
get_codes(9)["neonatal", ]
```

## ICD-10
```{r}
get_codes(10)["neonatal", ]
```

# Tech Dep
## ICD-9
```{r}
get_codes(9)["tech_dep", ]
```

## ICD-10
```{r}
get_codes(10)["tech_dep", ]
```

# Transplant
## ICD-9
```{r}
get_codes(9)["transplant", ]
```

## ICD-10
```{r}
get_codes(10)["transplant", ]
```

