---
title: "PCCC: An Example Using the Center for Disease Control's Multiple Cause of Death Dataset"
author: 
- Peter DeWitt
- James Feinstein
- Tell Bennett
- Seth Russell
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pccc-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Introduction
This vignette provides an example using actual publically available data to illustrate how the `pccc` package generates the Complex Chronic Condition Categories (CCC) from ICD-9 and ICD-10 codes. For an overview of the CCC classification system is avaiable, see [pccc-icd-overview](pccc-overview.html).

To evaluate the code chunks in this example you will need to load the
following R packages.

```{r, message = FALSE}
library(pccc)
library(dplyr)
```

# Accessing and Preparing the Data

* Will add language here about accessing the CDC's Multiple Cause of Death Data.
* Will add language here about preapring/processing the ICD9 and ICD10 codes for the analysis.

# Running the PCCC Software

* Will add language here about running PCCC software for both diagnosis and procedure codes.
* Will add chunks of corresponding code with tabulated CCC's.

# References

* Feudtner C, et al. Pediatric complex chronic conditions classification system version 2: updated for ICD-10 and complex medical technology dependence and transplantation, BMC Pediatrics, 2014, 14:199, DOI: 10.1186/1471-2431-14-199
* Feudtner C, et al. Pediatric deaths attributable to complex chronic conditions: a population-based study of Washington State, 1980-1997. Pediatrics. 2000;106(1 Pt 2):205-209.
