---
title: "Pediatric Complex Chronic Conditions"
author: 
- Peter DeWitt
- James Feinstein
- Tell Bennett
- Seth Russell
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pccc-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Introduction
This vignette describes how the `pccc` package generates the Complex Chronic Condition Categories (CCC) from ICD-9 and ICD-10 codes. 

A CCC is “any medical condition that can be reasonably expected to last at least 12 months (unless death intervenes) and to involve either several different organ systems or 1 organ system severely enough to require specialty pediatric care and probably some period of hospitalization in a tertiary care center." The categorization is based on the work of Feudtner et al. (2000 & 2014), as referenced below. 

A supplemental reference document showing the lists of codes for each category was published as a supplement to Feudtner et al. (2014) and we have made it available as part of the `pccc` package. After installing the package, you can find the file on your system with the below `system.file` call. Open the file with your preferred/available program for `.docx` files (Word, etc.).

```{r, eval = FALSE}
system.file("pccc_references/Categories_of_CCCv2_and_Corresponding_ICD.docx", package = "pccc")
```

To evaluate the code chunks in this example you will need to load the
following R packages.

```{r, message = FALSE}
library(pccc)
library(dplyr)
```

# Logic employed

There are 12 total categories of CCCs used in this package. The first group of 10 are mutually exclusive - only one of them can be derived from a ICD code:

* Neurologic and Neuromuscular
* Cardiovascular
* Respiratory
* Renal and Urologic
* Gastrointestinal
* Hematologic or immunologic
* Metabolic
* Other Congenital or Genetic Defect
* Malignancy
* Premature and Neonatal

The last 2 can be be selected in addition to the above codes - so for example, one ICD code could be both a Gastrointestinal and Technical Dependency:

* Technology Dependency
* Transplant

To see actual specific ICD codes by category, see [pccc-icd-codes](pccc-icd-codes.html)

# Generating CCC categories from ICD codes
The `ccc` function is the workhorse here. Simply put, a user will provide ICD codes as strings
and `ccc` will return CCC categories.

**Users of the `pccc` package will need to pre-process the ICD-9 and ICD-10 codes in their data so that the strings are formatted in the way that the `pccc` package will recognize them.** 

The ICD codes are matched exactly and should include leading zeroes for any ICD-9-CM codes (see specific rules below), but should not include periods.For example, ICD-9-CM 04.92 will only be matched by the string "0492". "492" _will not_ match.

Specific rules for ICD Codes are:

* Codes should be alphanumeric only - no periods, spaces or other separator characters (e.g. _Diabetes with renal manifestations, type II or unspecified type, uncontrolled_ should be sent as 25042)
* ICD 9 codes should be at a minimum 3 digits long:
    * Codes less than 10 should be left padded with 2 zeros. E.g. _Cholera due to vibrio cholerae el tor_ should be sent as 0011) 
    * Codes less than 100 should be left padded with 1 zero. E.g. _Whooping cough_ should be sent as 033)
    
Users of PCCC may find the R Package [ICD](https://jackwasey.github.io/icd/) useful.

# PCCC Examples

To illustrate the how the input formatting impacts the identification of a CCC, consider the data `data.frame` named `dat` below. These data have information about three patients (A-C). Each subject has the same ICD-9-CM diagnosis code (e.g. _Hypertrophic obstructive cardiomyopathy_, which should be sent as 4251) and the same ICD-9-CM procedure code (e.g. _Heart transplantation_, which should be sent as 3751), but each input is formatted differently. Based on the ICD-9-CM diagosis code, the `ccc` function will only identify subject `A` as having a CCC. Based on the ICD-9-CM procedure code, the `ccc` function will only identify subject `B` as having a CCC and will also flag the Transplantation category.

## Basic Example

```{r}
dat <- data.frame(ids = c("A", "B", "C"), 
                  dxs = c("4251", "425.1", "425.1"), 
                  procs = c("37.51", "3751", "37.51"))
dat
ccc(dat, 
    id = ids, 
    dx_cols = dxs, 
    pc_cols = procs, 
    icdv = 9)
```

## Extended Example

This example used a tool developed by Seth Russell (available at [icd_file_generator](https://github.com/magic-lantern/icd_file_generator)) to create a sample data file for ICD9 and ICD10. The generated data files contain randomly generated ICD codes for 1,000 patients and is comprised of 10 columns of diagnosis codes (d_cols), 10 columns of procedure codes (p_cols), and 10 columns of other data (g_cols).

Sample of how ICD9 test file was generated:

```{r eval = FALSE}
pccc_icd9_dataset <- generate_sample(
  v = 9,
  n_rows = 10000,
  d_cols = 10,
  p_cols = 10,
  g_cols = 10
)

save(pccc_icd9_dataset, file="pccc_icd9_dataset.rda")
```

Example using sample patient data set:

```{r}
library(dplyr)
library(pccc)

ccc_result <-
    ccc(pccc::pccc_icd9_dataset[, c(1:21)], # get id, dx, and pc columns
        id      = id,
        dx_cols = dplyr::starts_with("dx"),
        pc_cols = dplyr::starts_with("pc"),
        icdv    = 09)

# review results
head(ccc_result)

# view number of patients with each CCC
sum_results <- dplyr::summarize_at(ccc_result, vars(-id), sum) %>% print.data.frame

# view percent of total population with each CCC
dplyr::summarize_at(ccc_result, vars(-id), mean) %>% print.data.frame


```

# References

* Feudtner C, et al. Pediatric complex chronic conditions classification system version 2: updated for ICD-10 and complex medical technology dependence and transplantation, BMC Pediatrics, 2014, 14:199, DOI: 10.1186/1471-2431-14-199
* Feudtner C, et al. Pediatric deaths attributable to complex chronic conditions: a population-based study of Washington State, 1980-1997. Pediatrics. 2000;106(1 Pt 2):205-209.
